---
title: "Evaluate the ranking performance"
output: html_notebook
author: "Wytze Vlietstra"
date: '`r format(Sys.Date(), "%d-%m-%Y")`'
---

#TODO: Aggregeren op SNPs, aangezien die soms meerdere genes targeten. Daar apart de MAP voor berekenen.

# Introduction

Here, we evaluate how our gene candidate prioritization method performs. 
We compare its performance to two baselines: 1) when a random gene is chosen, and 2) when the closest gene is chosen.
Because a single SNP can target multiple genes, we use the mean average precision as an evaluation metric to compare the rankings.

# Transform the predictions to the proper format

```{r}
require(rjson)
load("Model")
candidates = read.csv2("Raw data files/Genes and gene candidates identified on 19-06-2019.csv", stringsAsFactors = F)
features = read.csv("Raw data files/Complete feature set generated on 17-06-2019.csv", stringsAsFactors = F)

scores = aggregate(list(score = model$pred$Target), by = list(rowIndex = model$pred$rowIndex), mean)
scores = scores[order(scores$rowIndex), -which(colnames(scores) == "rowIndex")]
scores = data.frame(Gene = features$Name, score = scores)

snps = unique(candidates$SNP.ID)
predictions = list()
for(i in seq_along(snps)){
  snp_candidates = unique(unlist(lapply(candidates$Candidate_genes[candidates$SNP.ID == snps[i]], fromJSON)))
  ranked_genes = scores[scores$Gene %in% snp_candidates, ]
  ranked_genes = ranked_genes[order(-ranked_genes$score),]
  ranked_genes$Class = ifelse(ranked_genes$Gene %in% candidates$Target_Gene[candidates$SNP.ID == snps[i]], "Target", "NonTarget")
  predictions[[snps[i]]] = ranked_genes
}
```

# Set baselines

First, we calculate the baseline performances against which we compare our performance.

## Random

```{r random}
sampling_repeats = 100
mean_sampled_ranks = rep(NA, times = length(predictions))
for(i in seq_along(predictions)){
  sampled_ranks = c()
  if("Target" %in% predictions[[i]]$Class){
    for(j in 1:sampling_repeats){
      sampled_ranks = append(sampled_ranks, sample(seq(1, nrow(predictions[[i]])), size = length(which(predictions[[i]]$Class == "Target"))))
    }
    mean_sampled_ranks[i] = mean(sampled_ranks)
  }
}
mean(na.omit(mean_sampled_ranks))
```

If we would randomly order the genes, we would achieve a mean average precision of.  

## Closest

```{r closest}


```

# Evaluate our predictions

```{r}
ranks = c()
for(i in seq_along(predictions)){
  if("Target" %in% predictions[[i]]$Class){
    ranks = append(ranks, which(predictions[[i]]$Class == "Target"))
  }
}

```

# Conclusion

# Used packages and R-version

```{r, echo = F}
sessionInfo()
```