---
title: "Distances between the SNPs and the genes that they target"
author: "Wytze Vlietstra"
date: '`r format(Sys.Date(), "%d-%m-%Y")`'
output:
  html_document: default
  html_notebook: default
---

```{r, echo = F, message = F, results = "hide"}
require(knitr)
options(scipen = 999)
```

# Introduction

Analysis of the input set.

Questions:

* How many of the targeted genes are not the closest gene
* What is the average distance of these genes?

# Load the data

```{r}
snps = read.csv2("Raw data files/Genes and gene candidates identified on 31-10-2019.csv", stringsAsFactors = F)

# Select only genes that can actually be found
snps = snps[snps$TargetGeneOnChromosome == T, ]
```

# Ranks

Based on distance alone, the target genes are located at the following ranks:

`r kable(table(snps$DistanceRankTargetGene, dnn = list("Rank")))`

# Distances

First, I show a histogram of all distances.

```{r, echo = F}
hist(log10(snps$DistanceToTargetGene), main = "Histogram of the distance of the SNPS to their target gene", xlab = "Number of basepairs distance in log 10", breaks = 16)
```

The different values are shown at different cutoffs:

`r kable(round(quantile(snps$DistanceToTargetGene, probs = seq(0, 1, 0.1))), col.names = "Distance")`

The mean difference is `r round(mean(snps$DistanceToTargetGene))`.

Next, I show a histogram of the distances for all genes that are not the closest gene.

```{r, echo = F}
hist(log10(snps$DistanceToTargetGene[snps$DistanceRankTargetGene > 1]), main = "Histogram of the distance of the SNPS to their target gene\nclosest gene excluded", xlab = "Number of basepairs distance in log 10", breaks = 10)
```

The different values are shown at different cutoffs:

`r kable(round(quantile(snps$DistanceToTargetGene[snps$DistanceRankTargetGene > 1], probs = seq(0, 1, 0.1))), col.names = "Distance")`

The mean difference is `r round(mean(snps$DistanceToTargetGene[snps$DistanceRankTargetGene > 1]))`.

# Suggested distances:

Based on the results above, I suggest the following distances to experiment with:

* 10000
* 25000
* 100000
* 250000
* 500000
* 1000000
* 1500000
* 2000000