---
title: "Machine learning experiments"
output: html_notebook
author: "Wytze Vlietstra"
date: '`r format(Sys.Date(), "%d-%m-%Y")`'
---

# Introduction

To identify genes that are likely targets of SNPs we train a machine learning classifier.
This notebook describes the experiments underlying the training and the analysis of the machine learning classifier.


# Load dataset and required packages

```{r package load, warning = F, message = F, results = "hide"}
require(yaml)
require(data.table)
require(caret)
require(ranger)
require(e1071)
require(pROC)
require(PRROC)
require(doMC)

config = read_yaml("Execution_config.yml")
registerDoMC(cores = config$coreCount)
```

```{r dataload}
removalColumns = c("EKP_ID", "EKP_Name", "Name", "Gene.start..bp.", "Gene.end..bp.",
                   "Chromosome.scaffold.name", "FC", "FDR", "expression", "Class")

f = as.data.frame(fread("Raw data files/Complete feature set generated on 2019-06-12.csv"))
outcome = factor(f$Class, levels = c("Target", "NonTarget"))

# Split up the dataset into features and non-features
meta_data = f[,which(colnames(f) %in% removalColumns)]
f = f[,which(!colnames(f) %in% removalColumns)]
```

Remove all infrequently occurring features
```{r nonZero}
nonZero = colSums(f != 0)
f = f[,names(nonZero[nonZero > 4])]
f2 = as.data.frame(f != 0)
f2[,c(1:18)] = f[,c(1:18)]
```

# Analysis of the input data

# Experiments

Set the general parameters for the machine learning algorithms that will be used in this document.
Only in specific experiments will the settings deviate.

```{r ML settings}
train_parameters = trainControl(method = "repeatedcv", number = 10, repeats = 10,
                                classProbs = TRUE, savePredictions = TRUE, 
                                summaryFunction = twoClassSummary, sampling = "down")
```

## Cross-validation performance
```{r rf, message = F, warning = F, results = "hide"}
model = train(y = outcome, x = f, method = "ranger", metric = "ROC", 
              tuneGrid = data.frame(mtry = round(sqrt(ncol(f))), splitrule = "gini", min.node.size = 1),
              save.memory = T,
              trControl = train_parameters, importance = "impurity", num.trees = 1000)
```

### Alternative performance metrics

Beside the ROC-AUC, we calculate several other performance metrics.
For the performance metrics that require a cut-off value, we determine the optimal cut-off by optimizing the sensitivity and the specificity in the ROC-curve.

```{r}
confMat = confusionMatrix(model$results$pred, model$results$obs)

```

### Distribution of prediction scores

### Distribution of misclassifications

## Iteratively remove the most predictive features

## Remove highly correlating features

## Different ratios of positive and negative cases in the training set

## Randomly assign outcome labels

## Other classifiers

### Linear regression

```{r glm, message = F, warning = F, results = "hide", cache = T}
model = train(y = outcome, x = f, method = "glm", metric = "ROC", 
              trControl = train_parameters)
```

### Neural network

```{r nnet, message = F, warning = F, results = "hide", cache = T}
model = train(y = outcome, x = f, method = "svmRadial", metric = "ROC", 
              trControl = train_parameters)
```

# Conclusions

# R-version and package versions

I used the following R-version and packages in these experiments:

```{r, echo = F}
sessionInfo()
```